---
title: "Case 3-1 Stat 440"
author: "Sarah Zimmermann, Adam Wood, Wuming Zhang"
date: "November 2, 2017"
output: pdf_document
---

#Introduction 

The data are from a study malaria presence in Gambia. Each observation corresponds to one child who has been tested for the presence of malaria. We are interested in the factors predictive of the malaria parasits found in the blood for n=805 children. The goal of the analysis is to perform inferences on the impact of age, presence of bednet, presence of clinic, and amount of surrounding greenry on a child having malaria, where greenery is a measure of how much vegetation is around the child's village based on satellite village. A summary of the variables is found below: 


Variable Name | Short Description | Type  
------------- | ------------- | --------------
Y|indicator of whether malaria parasites were found in the blood of the child|binary
AGE| age of child in years| continuous 
BEDNET| indicator of whether the child has a bednet over his or her bed| binary 
GREEN| a measure of how much greenery is around the child's village, derived from satellite images| continuous 
PHC| indicator for the presence of a public health clinic in the child's village| binary



#Data
```{r, warning=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)

library(ggplot2)
library(devtools)
library(ggplot2)
library(survival)
library(gridExtra)
library(scales)

gambia= read.csv("gambiaMissing.csv") 
dim(gambia)
```

#EDA

##Univariate Analysis
```{r}
#Univariate Analysis

pc1 = qplot(gambia$GREEN, geom = "histogram", main = "Histogram of Village Greenery", xlab = "Greenery Level", ylab = "Frequency", fill=I("blue"), binwidth = 0.5)
#overwhelming amount of data for villages with 40 value greenery levels. Very little data (could do percentage of total observations plot) for the below 40 or above 45 value greenery level villages


pc2 = qplot(gambia$PHC, geom = "histogram", main = "Histogram of Villages with Public Health Clinic", xlab = "Public Clinic Present", ylab = "Frequency", fill=I("blue"), binwidth = 0.5) 
#528 observations with a Public Clinic(65.6%) in the village, 277 without (34.4%)

pc3 = qplot(gambia$AGE, geom = "histogram", main = "Histogram of Child Age", xlab = "Child's Age", ylab = "Frequency", fill=I("blue"), binwidth = 0.5) 
#fairly even distribution for age of the child, with [29.6%, 23.4%, 24.7%, 22.4%] distribution of ages 
#could look into adding percentage of observations into the ggplot

pc4 = qplot(gambia$BEDNET, geom = "histogram", main = "Histogram of Bednet Use(Complete Cases)", xlab = "Bednet Presence", ylab = "Frequency", fill=I("blue"), binwidth = 0.5) 
#317 missing cases of Bednet use, 70.5% use a Bednet for complete cases while 29.5%

grid.arrange(pc1, pc2, pc3, pc4, ncol=2, nrow=2)
```
The histogram of village greenery illustrates that the majority of observations (72.8%) were in villages with a greenery value of 40.85. There is a relatively even representation amongst the other unique values of greenery, with greenery values of 28.85, 33.65, 40.10 and 47.65 recieving 3.4%, 10.4%, 8.6%, and 4.2% of observations, respectively. Conclusions and estimates of malaria presence in villages with greenery value other than 40.85 will have to carefully constructed. A heavily uneven distribution of this variable will make estimates about malaria presence in greeneries of values other than 40.85 tricky, and given the small volume of observations in the other unique values of greenery, high in uncertainty.

The distribution of the $PHC$ variable indicates that 528 observations (65.6%) have a Public Clinic in the village, while 277 observations are without a clinic (34.4%). The age of the children in this study is relatively evenly distributed, with ages 1 to 4 having a distribution of observations of [29.6%, 23.4%, 24.7%, 22.4%] respectively. An unevenly distributed age variable may lead to concerns about age bias in our conclusions, but even distribution of the variable means that each age will be appropriately represented. 

The bednet use variable has 317 missing cases, which will have to be handled in our analysis. Of the complete bednet cases in the data set, 70.5% use a Bednet while 29.5% do not. Bivariate analyses will provide more information about the relationships between these four variables and malaria presence. 


##Bivariate Analysis
```{r}
x= gambia[complete.cases(gambia), ]
 
x$Malaria[x$Y==1]= "Yes"
x$Malaria[x$Y==0]= "No"

x$Clinic[x$PHC==1]= "Yes"
x$Clinic[x$PHC==0]= "No"

x$Bednet[x$BEDNET==1]= "Yes"
x$Clinic[x$BEDNET==0]= "No"


p5<-ggplot(data = x, aes(x=BEDNET,fill=Malaria))+geom_bar(position="fill") +theme_bw()+ labs(x="Bednet",y="Percentage",fill="Presence of Malaira",title ="% of Ppl w/ Malaria by Bednet") + theme(plot.title = element_text(size=9, hjust = 0.5))+scale_y_continuous(labels=percent_format())


p4<-ggplot(data = x, aes(x=PHC,fill=Malaria))+geom_bar(position="fill") +theme_bw()+ labs(x="Clinic Present",y="Percentage",fill="Presence of Malaria",title ="% of Ppl w/ Malaria by Clinic Present") + theme(plot.title = element_text(size=9, hjust = 0.5))+scale_y_continuous(labels=percent_format())

p3<-ggplot(data = x, aes(x=AGE,fill=Malaria))+geom_histogram(position="fill", binwidth = .5) +theme_bw()+ labs(x="Age",y="Frequency",fill="Presence of Malaria",title ="% of Ppl w/ Malaria by Age") + theme(plot.title = element_text(size=9, hjust = 0.5))+scale_y_continuous(labels=percent_format())


p2<-ggplot(data = x, aes(x=GREEN,fill=Malaria))+geom_histogram(position="fill", binwidth = .5) +theme_bw()+ labs(x="Greenery",y="Frequency",fill="Presence of Malaria",title ="% of Ppl w/ Malaria by Greenery") + theme(plot.title = element_text(size=9, hjust = 0.5))+scale_y_continuous(labels=percent_format())


grid.arrange(p2, p3, p4,p5, ncol=2, nrow=2)
```
(Top Right) Here we visualize the percent of people with malaria based on the different greenery in the surrounding village. We see the percent of people with the disease varies with the amount of greenery. The relationship between the greenery and people sick is not a strictly positive or negative correlation because the percent of sick both increases and decreases as greenery increases. This is somewhat surprising as mosquitoes (which transmit the disease to humans) are likely to live in places with greenery; however, this graphic does not suggest the more greenery the higher the percent of children with malaria. 

(Top Left) Here we visualize the percent of people with and without malaria based across different ages. As age increases the percent of people in the age bracket sick increases. 

(Bottom Right) Here we visualize the percent of people with and without malaria based on the presence of a clinic in the village. A higher percent of people are sick in villages without a clinic compared to those sick in a village with a clinic. This makes sense that the presence of malaria is smaller where there are clinics because this means children have access to medicine. Additionally, it is possible the clinics probably help educate and spread awareness about prevention therefore decreasing the presence of malaria.   

(Bottom Left) Here we visualize the percent of people with and without malaria based on use of bednets. Of the child who do not have bednets there is not a significantly higher percent of children sick than those who used bed nets. This is surprising because bednets are a means of prevention and are recommended by agencies such as the CDC (https://www.cdc.gov/malaria/malaria_worldwide/reduction/itn.html). The data suggests, however, there is not much difference between the percent of people sick with and without bed use.

# Approaches Outline
## Missing Data Mechanisms
From initial exploration, we observe that we are missing 39.38% data for the explanatory variable BEDNET among all observations. We first suspect that this might simply be some random data collection issue. For instance, it could be that it is generally hard to collect data for people and missing data would occur very randomly; that is, the missing completely at random (MCAR) assumption. However, it is a bit suspicious that we have no missing data for any other explanatory variable in the dataset. Therefore, another possibility is that, it is harder to collect the BEDNET data for certain group of the people in this study than the others; in another word, the missing at random (MAR) assumption. 

To further explore whether certain groups are more likely to have missing value, we perform a logistic regression with missing value as th reponse variable and all other variables in the original dataset as the explanatory variables. We observe from the results below that, the coefficients for AGE and PHC variables are statistically significant. This means that whether missing data would occur has some correlation with some variables in the dataset. Therefore, we can reject the MCAR assumption and assume that only MAR assumption holds.
```{r}
#variable summaries 
#apply(gambia, 2, function(x) summary(as.factor(x)))


#missmech for MCAR assumption

#MAR
gambia$missing= ifelse(is.na(gambia$BEDNET), 1, 0)
summary(glm(missing~ Y+ AGE + GREEN+ as.factor(PHC), data=gambia ))

#Impute 
```

## Method proposal
A naive way of dealing with missing data is to perform listwise deletion, that is, to delete the entire observation if a missing data occurs. Although this method is very easy to implement, it would potentially reduce the statistical power of our analysis since we would need to reduce the size of the dataset by almost half. Additionaly, since the MCAR assumption does not hold, our estimates would be biased using this method. 

In order to obtain unbaised estimates while properly using the MAR assumption, we plan to use the multiple imputation method. This method contains three main steps. First, we impute the missing values for $m$ times using an appropriate model. Second, we perform our desired analysis on each of the $m$ datasets. Last, we combine the $m$ analysis into one final result using some proper method. 
Note that the multiple imputation method would bring us more variability thus improving the accuracy of our estimates. However, the implementation of this method could be much more challenging and we need to be very careful when choosing models. 


#Credit

Sarah wrote the introduction and created/analyzed bivariate graphs. Wuming wrote about missing data mechanisms and method proposal. 

#References
http://www.stat.columbia.edu/~gelman/arm/missing.pdf

http://www.statisticssolutions.com/multiple-imputation-for-missing-data/

https://liberalarts.utexas.edu/prc/_files/cs/Missing-Data.pdf


